---
title: "object-state-changes-and squashability"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---


#compare number of columns: 30 vs 31
ncol(prolific1)
ncol(prolific2)
ncol(sona_data)
colnames(prolific1)
colnames(prolific2)
colnames(sona_data)
#delete extra column in sona
sona_data <- sona_data %>% select(-c(sona_id))
```

# bind Prolific + Sona
```{r}
final_data = rbind(prolific1, prolific2, sona_data)
```

# basic descriptives
```{r}
final_data %>% pull(subject) %>% unique()
## N(initial) = 105
## note: R counts 'NA' as a participant 
final_data %>% group_by(subject) %>% count()
sona_data %>% pull(subject) %>% unique()
```

# exclusions
```{r}
##exclude filler, practice, and non-picture trials
data <- final_data %>% filter(typeoftrial=="picture" & !(trialcondition %in% c("filler", "practice")))
##calculate subject accuracy
data <- data %>% mutate(accuracy=ifelse(correct=="TRUE",1,0))
subject_accuracy <- data %>% group_by(subject) %>% summarise(TaskAccuracy=mean(accuracy))
##exclude incorrect responses
data <- data %>% filter(accuracy==1) 
##exclude subjects with accuracy > 60%
InaccurateSubjects <- subject_accuracy %>% filter(TaskAccuracy < 0.6)
data <- data %>% filter(!subject %in% InaccurateSubjects$subject)
##total participants after exclusion: N = 103 (2 excluded--33%, 58% accuracy)
length(data %>% pull(subject) %>% unique())
```

# transforming RT
```{r}
##make rt numeric
data <- data %>% mutate(rt = as.numeric(rt))
## rt histogram ----does NOT look normally distributed (positive skew)
data %>% filter(rt <10000) %>% ggplot(aes(x=rt)) + geom_histogram(binwidth =20, fill="pink" ) + theme_economist() + labs(x="RT (ms)", y="frequency", title= "RT histogram")
##rt log (base 10) transformation
data <- data %>% mutate(log10_rt=log(rt, 10))
##log10(rt) histogram
data %>% ggplot(aes(x=log10_rt)) + geom_histogram(binwidth =0.1, fill="pink" ) + theme_economist() + labs(x="log10(rt)", y="frequency", title= "RT histogram")
##looks fairly normal now
```

### MAD for individual trials by condition
```{r}
##removes 130/2275 trials
MADtable <- data %>% select(c(7,9,23,28,32))
trial_table <- MADtable %>% unite("trial_id", 1:2) %>% select(!rt)
a = trial_table %>% filter(trialcondition =="Squashable_Normal") %>% mutate(MAD=3*mad(log10_rt), med=median(log10_rt))
b= trial_table %>% filter(trialcondition =="Unsquashable_Normal") %>% mutate(MAD=3*mad(log10_rt), med=median(log10_rt))
c= trial_table %>% filter(trialcondition =="Squashable_Smashed") %>% mutate(MAD=3*mad(log10_rt), med=median(log10_rt))
d= trial_table %>% filter(trialcondition =="Unsquashable_Smashed") %>% mutate(MAD=3*mad(log10_rt), med=median(log10_rt))
## Note these are all 3*MAD
a = a %>% mutate(L= (med - MAD), U=(med + MAD))
b=b %>% mutate(L= (med - MAD), U=(med + MAD))
c=c %>% mutate(L= (med - MAD), U=(med + MAD))
d=d %>% mutate(L= (med - MAD), U=(med + MAD))
data2 <- rbind(a,b,c,d)
outliers <- data2 %>% filter((log10_rt > U) | (log10_rt < L))
dataS <- data
data <- data %>% unite("trial_id", c(7,9), sep="_")
data <- data %>% filter(!trial_id %in% outliers$trial_id)
data <- data %>% separate(trial_id, into = c("subject", "trial_index"), sep="_")
```



# plot
```{r}
data = data %>% separate(trialcondition, into = c("squashability", "state")) 
data %>%
  group_by(squashability, state) %>%
   summarise(ci = list(mean_cl_boot(rt) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  ggplot(aes(x= squashability, y = mean, group = state, fill = state)) +
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "mean ratings", title = "squashability ratings")+
  theme_clean()+
  scale_fill_manual(values = c("darksalmon", "darkslategray4","darkolivegreen4"))+
  theme(plot.background = element_rect(
    color = "white"),
    strip.text.x = element_text(size =rel(2)),
        axis.text = element_text(size =rel(1.5)),
        axis.title = element_text(size =rel(2)),
        plot.title = element_text(hjust = .5, size = rel(1.5)))
```

```{r}
data %>% group_by(squashability, state) %>% shapiro_test(log10_rt)
##qqplot
ggqqplot(data, "log10_rt", ggtheme = theme_bw()) + facet_grid(squashability ~ state, labeller = "label_both")
```


# model
```{r}
library(lmerTest)
rt_model = lmer(data = data, log10_rt ~ squashability*state + (1|subject) + (1|stimulus))
summary(rt_model)
car::Anova(rt_model)
```

# assumptions
```{r}
library(glmmTMB)
sjPlot::plot_model(rt_model, type='diag')
sjPlot::plot_model(rt_model, type='resid')
```



# compute mean RTs
```{r}
##Here are each subjects' means for all 4 conditions, and then the overall means for each condition...
subjectMeans<- dataS %>% group_by(subject, trialcondition) %>% summarise(subj_mean=mean(log10_rt), subj_meanRaw=mean(rt)) 
conditionMeans <- dataS %>% group_by(trialcondition) %>% summarise(conditionMean=mean(log10_rt))
```

### MAD by subjects per condition
```{r}
##removes 33 subject trials
e = subjectMeans %>% ungroup() %>% filter(trialcondition =="Squashable_Normal") %>% mutate(MAD=3*mad(subj_mean, constant = 1), med=median(subj_mean))
f= subjectMeans %>% ungroup() %>% filter(trialcondition =="Unsquashable_Normal") %>% mutate(MAD=3*mad(subj_mean, constant = 1), med=median(subj_mean))
g= subjectMeans %>% ungroup() %>% filter(trialcondition =="Squashable_Smashed") %>% mutate(MAD=3*mad(subj_mean, constant = 1), med=median(subj_mean))
h= subjectMeans %>% ungroup() %>% filter(trialcondition =="Unsquashable_Smashed") %>% mutate(MAD=3*mad(subj_mean, constant = 1), med=median(subj_mean))
## Note these are all 3*MAD
e = e %>% mutate(L= (med - MAD), U=(med + MAD))
f=f %>% mutate(L= (med - MAD), U=(med + MAD))
g=g %>% mutate(L= (med - MAD), U=(med + MAD))
h=h %>% mutate(L= (med - MAD), U=(med + MAD))
subj_MADranges<- rbind(e,f,g,h)
outliers1 <- subj_MADranges %>% filter((subj_mean > U) | (subj_mean < L))
subjectMeans1 <- subjectMeans %>% filter(!subject %in% outliers1$subject)
```

# check assumptions
## identify and remove outliers
```{r}
##outliers <- subjectMeans %>% group_by(trialcondition) %>% identify_outliers(subj_mean)
##outlier_exclusion <- outliers %>% filter(is.extreme=="TRUE")
#filter extreme outliers
##subjectMeans1 <- subjectMeans %>% filter(!subject %in% outlier_exclusion$subject)
##filtering all outliers
##subjectMeans2 <- subjectMeans %>% filter(!subject %in% outlier_exclusion$subject)
subjectMeans1 %>% pull(subject) %>% unique()
##subjectMeans2 %>% pull(subject) %>% unique()
## 14 outliers, 3 extreme outliers, 3 exclusions
```
## check normality
```{r}
##shapiro test
subjectMeans1 %>% group_by(trialcondition) %>% shapiro_test(subj_mean)
##all have p>0.05, but squash-Smashed condition is close at p=0.052. Data satisfies normality condition?
##qqplot....plots look fine, not perfect 
subjectMeans1_split <- subjectMeans1 %>% separate(trialcondition, into=c("squashability","state"), sep="_")
ggqqplot(subjectMeans1_split, "subj_mean", ggtheme = theme_bw()) + facet_grid(squashability ~ state, labeller = "label_both")
```

# conduct repeated-measures anova
```{r}
subjectMeans1_split <- subjectMeans1_split %>% ungroup() 
RManova <- anova_test(data = subjectMeans1_split, dv=subj_mean, wid = subject, within = c(squashability, state), effect.size = "pes")
get_anova_table(RManova)
##no significant main effect for squashability
##significant main effect for state, significant interaction
##no sphericity violation
```

# conduct post-hoc tests (if needed)
```{r}
##decomposing significant interaction
subjectMeans1_split %>% group_by(squashability) %>% anova_test(dv=subj_mean, wid = subject, within = state, effect.size = "pes") %>% get_anova_table() %>% adjust_pvalue(method = "bonferroni")
#### squashable: p.adj = 1.000
#### unsquashable: p.adj = .00000155 *, partial eta-squared=0.207 (large effect)
#### The RT for the unsquashable objects was significantly different depending on the state of the target object (smashed vs normal), but there was no significant difference found for squashable objects when comparing thee reaction times for normal vs smashed targets. This pattern in the the graph below shows that RTs for unsquashable objects were significantly slower when the target object was smashed compared to when they were normal. This suggests that the participants activated the normal but not the smashed state for the target object when primed by the sentence, despite the fact that a bowling ball was being dropped on it, causing a slower RT than when the normal object was shown. For the squashable objects, the data interpreted in line with the Horachak & Garrido study suggests that both the smashed and unsmashed sates were activated when primed by the bowling ball sentences, resulting in no difference between reaction times when a smashed vs normal image of thw squashable object was shown.
```

# visualize log and raw RT pattern
```{r}
conditionMeans <- subjectMeans1 %>% group_by(trialcondition) %>% summarise(rawRTmean=mean(subj_meanRaw), sd_r=sd(subj_meanRaw), LogRTmean=mean(subj_mean), sd_L=sd(subj_mean))
## SEM calculations
n= subjectMeans1 %>% group_by(subject) %>% n_groups() 
conditionMeans <- conditionMeans %>% mutate(SEM_r=sd_r/sqrt(n), SEM_L=sd_L/sqrt(n))
##graphs
conditionMeans %>% separate(trialcondition, into = c("squashability","state"), sep="_") %>% ggplot(aes(x=squashability, y=rawRTmean, fill=state)) + geom_col(position = "dodge") +  labs(x="object squashability", y="RT (ms)", title = "RT interaction of object state and squashability") + geom_errorbar(aes(ymin= (rawRTmean - SEM_r), ymax= (rawRTmean + SEM_r)), width=.25, position=position_dodge(width=0.9)) + theme_economist_white() + scale_fill_economist()
   
conditionMeans %>% separate(trialcondition, into = c("squashability","state"), sep="_")  %>% ggplot(aes(x=squashability, y=LogRTmean, fill=state)) + geom_col(position = "dodge") + labs(x="object squashability", y="log10(RT)", title = "log10(RT) interaction of object state and squashability") + geom_errorbar(aes(ymin= (LogRTmean - SEM_L), ymax=(LogRTmean + SEM_L)), width=.25, position=position_dodge(width=0.9))  + theme_economist_white() + scale_fill_economist()
   
```
## Demographic Info
```{r}
d = final_data
demo = d %>% filter(typeoftrial %in% c("demo", "english") & response != "0") %>%
  separate(response, into = c("demo1", "demo2", "demo3", "demo4"), sep = ":")
age = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo2) %>%
  mutate(age = as.numeric(gsub("[^[:digit:] ]", "", demo2)))
age %>% summarise(mean_age = mean(age))
race = demo %>% filter(demo1 == "{\"Race\"") %>% select(demo2)
race %>% group_by(demo2) %>% count()
gender_education = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo3) %>%
  separate(demo3, into = c("gender", "education"), sep = ",")  %>%
  mutate(gender = tolower(gender),
         gender = gsub("[[:space:]]", "", gender))
gender_education %>% group_by(gender) %>%
  count()
gender_education %>% group_by(education) %>%
  count()
first_lang = demo %>% filter(demo1 == "{\"English\"") %>% select(demo2)
first_lang %>% group_by(demo2) %>% count()
```

```{r}
"18-19" =(age %>% filter(c(age) <= 19 & c(age) >= 18) %>% count())
"20-29" = (age %>% filter(c(age) <= 29 & c(age) >= 20) %>% count())
"30-39" = (age %>% filter(c(age) <= 39 & c(age) >= 30) %>% count())
"40-49" = age %>% filter(c(age) <= 49 & c(age) >= 40) %>% count()
"50-59" = (age %>% filter(c(age) <= 59 & c(age) >= 50) %>% count())
"60+" = (age %>% filter(c(age) >= 60) %>% count())
```