---
title: "OS Survey Sanity Check"
author: "Channing Hambric"
date: "2025-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
library(dplyr)
library(ggthemes)
library(ggplot2)
library(stringr)
library(paletteer)
library(patchwork)

```

# Import Data
```{r}
#Data from pilot - N=2
data <-read.csv("../data/Object_State_Survey_Pilot_Testing_Prolific.csv", fileEncoding = "UTF-8") %>%rename("id"="PROLIFIC_PID")

```

# Qualtrics data cleaning & reorganization
```{r}
# Remove second row (assuming it's unwanted)
data <- data[-c(2), ]

# Select demographic columns
demos <- data %>%
  select(id,Age, Gender, Race, Hispanic, NES, EngAOA,NativeLang) %>%
  slice(-1)

# Move 'id' to the first column and delete demos
pre_clean_data <- data %>% relocate(id, .before = everything()) %>% select(-matches("Practice")) %>% select(-c(Age, Gender, Race, Hispanic, NES, NativeLang))


#Below will need to be changed for PG (adding null sub condition)
#Setting Up Rating Data
rating_data <- pre_clean_data %>%
  select(1, matches("_Before_|_After_|_Imply_")) %>%
  slice(-1) %>%
  pivot_longer(cols = -1,  
               names_to = "Condition",  
               values_to = "Score") %>%
  separate(Condition, 
           into = c("verb", "tense", "target_object", "question", "focus", "extra"), 
           sep = "_", 
           fill = "right") %>%
  select(-extra) %>%
  mutate(Score = na_if(Score, "")) %>%  
  mutate(Score = na_if(Score, " ")) %>% 
  drop_na(Score) %>%
  mutate(
    verb = recode(verb, "D" = "Drop", "C" = "Contact"),
    tense = recode(tense, "Pr" = "Present", "Pa" = "Past"),
    target_object = recode(target_object, "T" = "tomato", "B" = "Blueberry", "M" = "muffin"),
    focus = recode(focus, "NF" = "NotFocus", "F" = "Focus"),
    focus = recode(focus, "1" = "NA"),
    Score = as.numeric(Score)  # Convert Score to numeric scale
  )

#Setting Up Ranking Data
ranking_data <- pre_clean_data %>%
  select(1, matches("_Rank_"))  

# Extract ranked objects
new_names <-ranking_data %>%select(-1)%>%
  slice(1) %>%  # Select the first row
  unlist() %>%  # Convert to vector
  str_split(" - ", simplify = TRUE) %>%  # Split by " - "
  .[, 2]  # Get the second part after the hyphen

# Replace numeric suffix in column names with extracted objects
colnames(ranking_data)[-1] <- str_replace(colnames(ranking_data)[-1], "\\d+$", new_names)

# Separate out conditions
ranking_data <- ranking_data %>% 
  slice(-1) %>%
  pivot_longer(cols = -1,  
               names_to = "Condition",  
               values_to = "Ranking") %>% 
  mutate(Ranking = na_if(Ranking, "")) %>%  
  mutate(Ranking = na_if(Ranking, " ")) %>% 
  drop_na(Ranking) %>%
  mutate(Condition_type = ifelse(str_starts(Condition, "C"), "C", "D"))  # Identify if drop or contact verb

# Separate contact conditions
c_conditions <- ranking_data %>%
  filter(Condition_type == "C") %>%
  separate(Condition, 
           into = c("verb", "tense", "target_object", "question", "focus", "ranked_object"), 
           sep = "_", 
           fill = "right")

# Separate drop conditions
d_conditions <- ranking_data %>%
  filter(Condition_type == "D") %>%
  separate(Condition, 
           into = c("verb", "tense", "target_object", "question", "ranked_object"), 
           sep = "_", 
           fill = "right") %>%
  mutate(focus = NA)  # Set focus to NA for 'D' conditions

# Combine the results
ranking_data <- bind_rows(c_conditions, d_conditions) %>%
  select(-Condition_type) %>% 
  mutate(
    verb = recode(verb, "D" = "Drop", "C" = "Contact"),
    tense = recode(tense, "Pr" = "Present", "Pa" = "Past"),
    target_object = recode(target_object, "T" = "tomato", "B" = "Blueberry", "M" = "muffin"),
    focus = recode(focus, "NF" = "NotFocus", "F" = "Focus")
  ) %>% 
  filter(!is.na(Ranking) & str_trim(Ranking) != "") %>%
  mutate(ranked_object_recoded = recode(ranked_object,
                                "the muffin" = "target object",
                                "the tomato" = "target object",
                                "the blueberry" = "target object",
                                "the brick" = "falling object",
                                "the bowling ball" = "falling object"))  

```


##DESIGN:
[English]
6 conditions total, 4 questions per condition, each condition presented w/ 1 of 3 items [tomato,blueberry,muffin] 
Each condition: Questions 1-3 continuous numeric scale, Q4 rank order (1-3 or 1-2)

Conditions [English]
Drop_Present
Drop_Past
Contact_Present_NotFocus
Contact_Past_NotFocus
Contact_Present_Focus
Contact_Past_Focus

[Portuguese]
8 conditions total, 4 questions per condition, each condition presented w/ 1 of 3 items [tomato,blueberry,muffin] 
Each condition: Questions 1-3 continuous numeric scale, Q4 rank order (1-3 or 1-2)

Conditions [Portuguese]
Drop_Present_NullSubject
Drop_Past_NullSubject
Drop_Present_ExpSubject
Drop_Past_ExpSubject
Contact_Present_NotFocus
Contact_Past_NotFocus
Contact_Present_Focus
Contact_Past_Focus

## Count subjects
```{r}
#should be 2 in each
rating_data %>% group_by(id) %>% count()
ranking_data %>% group_by(id) %>% count()
```


## Ratings Check
```{r}
#Should be 6 conditions
rating_data %>% group_by(verb,tense,focus) %>% count()

#BEFORE Q- Should be 1 before question per participant per combined condition
rating_data %>% filter(question=="Before") %>%group_by(id,verb,tense,focus) %>% count()
#Should be 2 across 2 subs
rating_data %>% filter(question=="Before") %>%group_by(verb,tense,focus) %>% count()

#AFTER Q- Should be 1 before question per participant per combined condition
rating_data %>% filter(question=="After") %>%group_by(id,verb,tense,focus) %>% count()
#Should be 2 across 2 subs
rating_data %>% filter(question=="After") %>%group_by(verb,tense,focus) %>% count()

#IMPLY Q - Should be 1 before question per participant per combined condition
#BEFORE - Should be 1 before question per participant per combined condition
rating_data %>% filter(question=="Imply") %>%group_by(id,verb,tense,focus) %>% count()
#Should be 2 across 2 subs
rating_data %>% filter(question=="Imply") %>%group_by(verb,tense,focus) %>% count()

```


## Ranking Check
```{r}
#Should be 6 conditions
ranking_data %>% group_by(verb,tense,focus) %>% count()

#For each subject, should be 3 in Drop conditions (since "you" is also ranked)  and 2 in Contact conditions
ranking_data  %>%group_by(id,verb,tense,focus) %>% count()

#Should be 6 (3 obj per ranking * 2 sub) in each Drop and 4 (2 obj per ranking * 2) in each Contact, across 2 subs
ranking_data  %>%group_by(verb,tense,focus) %>% count()

```
## Demos check

```{r}
#Age, race, gender - should be 2 entries total
demos %>% group_by(Age,Race,Gender) %>% count()

#how many Nonnative Eng speakers, Nonnative English AOA - should add up to 2
demos%>%group_by(NES,NativeLang,EngAOA) %>% count()
```


