---
title: "object-state-changes-sentence-focus"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

# load packages
```{r}
library(tidyverse)
library(ggplot2)
library (ggthemes)
library(rstatix)
library(ggpubr)
library(emmeans)
library(dplyr)
library(lmerTest)
```

# import data
```{r}

pilot_data <- read_csv("../data/sona_pilot.csv")

```

# basic descriptives
```{r}
length(pilot_data %>% pull(subject) %>% unique())
## N(final) = 10
pilot_data %>% group_by(subject) %>% count()
```
# data treatment

## exclude non-English first language participants***
```{r}
lang <- pilot_data %>% filter(english=="no") 
pilot_data <- pilot_data %>% filter(!subject %in% lang$subject)
```

## exclusions
```{r}
##exclude filler, practice, and non-picture trials
exclusions_data <- pilot_data %>%
  filter(typeoftrial=="picture" & !(trialcondition %in% c("filler", "practice_s","practice_n"))) %>%
  ##calculate subject accuracy
  mutate(accuracy=ifelse(correct=="TRUE",1,0)) %>% 
  separate(trialcondition, into = c("weight", "state","focus"))
```

## item labels 
```{r}
items <- gsub('.{0,5}$', '', exclusions_data$object) %>% as.data.frame()
exclusions_data <- exclusions_data %>% mutate(item = items[,1] )

### For experimental picture trials: this code creates a new column 'items' to id the item (picture stimulus) used in each trial. The column 'stimulus' has the picture file name and specifies the smashed/intact picture state. This new col only has the item  (ex. the original stimulus label might be "tomatoS.png" or tomatoN.png" -> new label is "tomato")
```


# accuracy analyses

##accuracy data
```{r}
subject_accuracy <- exclusions_data %>% group_by(subject) %>% summarise(TaskAccuracy=mean(accuracy))
InaccurateSubjects <- subject_accuracy %>% filter(TaskAccuracy < 0.8)
```

## plot
```{r}
exclusions_data %>% 
  group_by(weight, state,focus) %>%
   summarise(ci = list(mean_cl_boot(accuracy) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  mutate(state = fct_recode(state, canonical = "Normal", `non-canonical` = "Smashed" ),
         state = fct_relevel(state, "non-canonical","canonical"))%>%
  ggplot(aes(x= state, y = mean, group = weight, fill = weight)) + facet_wrap(~focus)+
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "mean accuracy", title = "")+
  theme_clean()
```
## main accuracy model
```{r}
acc_model = glmer(data = exclusions_data, 
                  accuracy ~ state*weight*focus + 
                    (1 + state + weight + focus|subject) + (1|stimulus),
                  family = "binomial", 
                 control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(acc_model)
car::Anova(acc_model)

sjPlot::plot_model(acc_model, type = "int")
```


# RT analyses

## rt data
```{r}
##exclude incorrect responses
rt_data <- exclusions_data %>% filter(accuracy==1) %>%
  ##exclude subjects with accuracy < 80%
  filter(!subject %in% InaccurateSubjects$subject)

##total participants after exclusion: N = 113 (5 excluded)
length(rt_data %>% pull(subject) %>% unique())
```

## transforming RT
```{r}
##make rt numeric
rt_data <- rt_data %>% mutate(rt = as.numeric(rt))
## rt histogram ----does NOT look normally distributed (positive skew)
rt_data %>% filter(rt <10000) %>% 
  ggplot(aes(x=rt)) + geom_histogram(binwidth =20, fill="pink" ) + 
  theme_economist() + labs(x="RT (ms)", y="frequency", title= "RT histogram")

##rt log (base 10) transformation
rt_data <- rt_data %>% mutate(log10_rt=log(rt, 10))

##log10(rt) histogram
rt_data %>% 
  ggplot(aes(x=log10_rt)) + geom_histogram(binwidth =0.1, fill="pink" ) + 
  theme_economist() + labs(x="log10(rt)", y="frequency", title= "RT histogram")
```

## MAD exclusions
```{r}
mad_data = rt_data %>%
  group_by(weight, state,focus) %>%
  mutate(MAD=3*mad(log10_rt), med=median(log10_rt),
         L= (med - MAD), U=(med + MAD)) %>%
  filter(log10_rt < U) %>% filter(log10_rt > L)

(nrow(rt_data)-nrow(mad_data))/nrow(rt_data)
```

## plot
```{r}
mad_data %>% 
  group_by(weight, state,focus) %>%
   summarise(ci = list(mean_cl_boot(rt) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  mutate(state = fct_recode(state, canonical = "Normal", `non-canonical` = "Smashed" ),
         state = fct_relevel(state, "non-canonical","canonical"))%>%
  ggplot(aes(x= state, y = mean, group = weight, fill = weight)) + facet_wrap(~focus)+
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "RT (ms)", title = "")+
  theme_clean()
```

## main RT model
```{r}
rt_model = lmer(data = mad_data, 
                log10_rt ~ weight*state*focus +
                  (1 + weight+state+focus |subject) + (1|stimulus),
                REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
fixef(rt_model)
summary(rt_model)
car::Anova(rt_model)
```
### power analyses

```{r}
library(simr)
# Start with an extended model up to 100 participants
fixef(rt_model)["weightLight:stateSmashed:focusNotFocus"] = 0.05

extended_model <- extend(rt_model, along = "subject", n = 250)
# Perform a power curve analysis to test the three-way interaction
power_curve <- powerCurve(extended_model, 
                          along = "subject", 
                          fixed("weight:state:focus", "anova"), 
                          nsim = 100)  # Number of simulations per sample size

# Plot the power curve
plot(power_curve)
```

### assumptions
```{r}
library(glmmTMB)
sjPlot::plot_model(rt_model, type='diag')
sjPlot::plot_model(rt_model, type='resid')
```

### posthoc models 

```{r}
# split into picture type (state)
intact_data = mad_data %>% filter(state == "Normal")
squashed_data = mad_data %>% filter(state == "Smashed")
# run separate models on these data
intact_model = lmer(data = intact_data, log10_rt ~ weight*focus + (weight+focus | subject) + 
                      (1 | item), REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
                  
summary(intact_model)
car::Anova(intact_model)


squashed_model = lmer(data = squashed_data, log10_rt ~ weight*focus + (weight+focus | subject) + 
                      (1 | item), REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(squashed_model)
car::Anova(squashed_model)
```

# Demographics and other Descriptives

## Experiment Duration (minutes)
```{r}
duration <- pilot_data %>% filter(typeoftrial == "english") %>%
  mutate(time_elapsed = time_elapsed/1000/60)

duration %>% pull(time_elapsed) %>% hist(40)
duration %>% pull(time_elapsed) %>% boxplot()

duration %>%  summarise(med = median(time_elapsed), iqr = IQR(time_elapsed), max = max(time_elapsed), min = min(time_elapsed))
quantile(duration$time_elapsed, c(.25,.75))
```

## Demographic Info
```{r}
demo = pilot_data %>% filter(typeoftrial %in% c("demo") & response != "0") %>%
  separate(response, into = c("demo1", "demo2", "demo3", "demo4","demo5"), sep = ":")

age <- demo[!is.na(demo$age), ] %>% summarise(mean_age = mean(age))

race = demo %>% filter(demo1 == "{\"Race\"") %>% select(demo2)
race %>% group_by(demo2) %>% count()

gender_education = demo %>% select(gender,education) %>%na.omit()
gender_education %>% group_by(gender) %>%
  count()
gender_education %>% group_by(education) %>%
  count()


#change above to save full part data in other var name
first_lang = demo %>% filter(demo1 == "{\"English\"") %>% select(demo2)
first_lang %>% group_by(demo2) %>% count()
```
