---
title: "slicing experiment"
output:
  word_document: default
  html_document:
    df_print: paged
---

# load packages
```{r}
library(tidyverse)
library(ggplot2)
library (ggthemes)
library(rstatix)
library(ggpubr)
library(emmeans)
library(dplyr)
library(paletteer)
library(lmerTest)
library(patchwork)
library(optimx)
setwd(here::here())
getwd()
```

# import & bind data
```{r}
## read sona and prolific data

prolific_data <- read_csv('../data/prolific_data.csv')
sona_data <- read_csv('../data/sona_data.csv') 
prolific_data_2 <- read_csv('../data/slicing_data2.csv')

## remove extra col.s and problematic col.
sona_data <- sona_data %>% select(!c(sona_id, subject_id, recorded_at))
prolific_data <- prolific_data %>% select(!c(subject_id, recorded_at))
prolific_data_2 <- prolific_data_2 %>% select(!c(recorded_at))

##bind prolific & sona data
data <- rbind(prolific_data, sona_data, prolific_data_2) %>%
  filter(!is.na(subject))
```

# basic descriptives and first langugae check
```{r}
## N(total) = 106
length(data %>% pull(subject) %>% unique())
```

# data treatment
## exclude non-English first language participants***
```{r}
lang <- data %>% filter(typeoftrial == "notenglish") %>% 
  separate(response, into = c("first", "age"), sep = ",") %>% 
  separate(first, into = c("q0", "first"), sep = ":") %>% mutate(first = gsub(" ", "", first)) %>%
  filter(toupper(first) != '"ENGLISH"')

data <- data %>% filter(!subject %in% lang$subject)
length(data %>% pull(subject) %>% unique())
```

## exclusions
```{r}
##exclude filler, practice, and non-picture trials
exclusions_data <- data %>%
  filter(typeoftrial=="picture" & !(trialcondition %in% c("filler", "practice"))) %>%
  ##calculate subject accuracy
  mutate(accuracy=ifelse(correct=="TRUE",1,0)) %>% 
  separate(trialcondition, into = c("verb", "state"))
```

# accuracy analyses

##accuracy data
```{r}
subject_accuracy <- exclusions_data %>% group_by(subject) %>% summarise(TaskAccuracy=mean(accuracy))
InaccurateSubjects <- subject_accuracy %>% filter(TaskAccuracy < 0.8)
exclusions_data %>% summarise(overall.acc = mean(correct))
```

## plot
```{r}
accuracy_plot <- exclusions_data %>% 
  group_by(verb, state) %>%
  summarise(ci = list(mean_cl_boot(accuracy) %>% 
                        rename(mean = y, lwr = ymin, upr = ymax))) %>% 
  unnest(ci) %>%
  mutate(state = fct_recode(state, `canonical` = "whole", `sliced` = "slice"),
         state = fct_relevel(state, "canonical", "sliced")) %>%
  ggplot(aes(x = state, y = mean, group = verb, fill = verb)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24") +
  geom_errorbar(aes(ymin = lwr, ymax = upr), size = 0.5, width = .15, 
                color = "black", position = position_dodge(0.70)) +
  labs(y = "accuracy", title = "") +
  theme_clean() +
  scale_fill_paletteer_d("ggthemes::Classic_10_Light") +
  theme(
    legend.position = "none",
    aspect.ratio = 1.5,
    plot.background = element_rect(color = "white"),
    strip.text.x = element_text(size = rel(2.3)),
    axis.text = element_text(size = rel(2.3)),
    axis.title = element_text(size = rel(3)),
    plot.title = element_text(hjust = 0.5, size = rel(2)),
    legend.text = element_text(size = rel(2)),  
    legend.title = element_text(size = rel(2))
  )

accuracy_plot

```
## main accuracy model
```{r}
acc_model_0 = glmer(data = exclusions_data, 
                  accuracy ~ state*verb + (1|subject) + (1|stimulus),
                  family = "binomial", 
                 control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))

acc_model = glmer(data = exclusions_data, 
                  accuracy ~ state*verb + (1 + state*verb|subject) + (1|stimulus),
                  family = "binomial", 
                 control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(acc_model)

performance::icc(acc_model)
anova(acc_model_0,acc_model)


```

No interaction but trending effects.

# RT analyses

## rt data
```{r}
##exclude incorrect responses
rt_data <- exclusions_data %>% filter(accuracy==1) %>%
  ##exclude subjects with accuracy < 80%
  filter(!subject %in% InaccurateSubjects$subject)

##total participants after exclusion: N = 89 
length(rt_data %>% pull(subject) %>% unique())
```


## transforming RT
```{r}
##make rt numeric
rt_data <- rt_data %>% mutate(rt = as.numeric(rt))
## rt histogram ----does NOT look normally distributed (positive skew)
rt_data %>% filter(rt <10000) %>% 
  ggplot(aes(x=rt)) + geom_histogram(binwidth =20, fill="pink" ) + 
  theme_economist() + labs(x="RT (ms)", y="frequency", title= "RT histogram")

##rt log (base 10) transformation
rt_data <- rt_data %>% mutate(log10_rt=log(rt, 10))

##log10(rt) histogram
rt_data %>% 
  ggplot(aes(x=log10_rt)) + geom_histogram(binwidth =0.1, fill="pink" ) + 
  theme_economist() + labs(x="log10(rt)", y="frequency", title= "RT histogram")
```

## MAD exclusions
```{r}
mad_data = rt_data %>%
  group_by(verb, state) %>%
  mutate(MAD=3*mad(log10_rt), med=median(log10_rt),
         L= (med - MAD), U=(med + MAD)) %>%
  filter(log10_rt < U) %>% filter(log10_rt > L)

(nrow(rt_data)-nrow(mad_data))/nrow(rt_data)
```

## plot
```{r}
rt_plot<-mad_data %>% 
  group_by(verb, state) %>%
   summarise(ci = list(mean_cl_boot(rt) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest %>%
  mutate(state = fct_recode(state, canonical = "whole", `sliced` = "slice" ),
         state = fct_relevel(state, "canonical","sliced"))%>%
  ggplot(aes(x= state, y = mean, group = verb, fill = verb)) +
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "response time (ms)", title = "",x = "state",
       fill = "sentence type") +
  theme_clean()+
   scale_fill_paletteer_d("ggthemes::Classic_10_Light")+
  theme(
    legend.position = "top",
    aspect.ratio = 1.5,
    plot.background = element_rect(color = "white"),
    strip.text.x = element_text(size = rel(2.3)),
    axis.text = element_text(size = rel(2.3)),
    axis.title = element_text(size = rel(3)),
    plot.title = element_text(hjust = 0.5, size = rel(2)),
    legend.text = element_text(size = rel(2)),  
    legend.title = element_text(size = rel(2)))

accuracy_plot+rt_plot
```

## main RT model
```{r}

rt_model = lmer(data = mad_data, 
                log10_rt ~ verb*state +
                  (1 + verb*state |subject) + (1|stimulus),
                REML = FALSE, 
                control = lmerControl(optimizer ='optimx', optCtrl=list(method='nlminb'),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(rt_model)

emmeans(rt_model,pairwise~verb*state, simple = "verb")
```

Significant interaction.

## power analysis
```{r}
library(simr)
#set.seed(123)
#Change the fixed effect to be small
#fixef(rt_model)["verbneutral:statewhole"] = 0.05
#Start with an extended model up to 100 participants
#extended_model <- extend(rt_model, along = "subject", n = 200)
#Perform a power curve analysis to test the three-way interaction
#power_curve <- powerCurve(extended_model, 
                        #  along = "subject", 
                        #fixed("verb:state", "anova"), 
                       #  nsim = 100)  # Number of simulations per sample size

# Plot the power curve
#plot(power_curve)

```

### assumptions
```{r}
library(glmmTMB)
sjPlot::plot_model(rt_model, type='diag')
sjPlot::plot_model(rt_model, type='resid')
```

### posthoc models 

```{r}
# split into picture type (state)
#intact_data = mad_data %>% filter(state == "whole")
#squashed_data = mad_data %>% filter(state == "slice")
# run separate models on these data
#intact_model = lmer(data = intact_data, log10_rt ~ verb + (verb | subject) + 
#                      (1 | stimulus), REML = FALSE, 
#                lmerControl(optimizer = "bobyqa",
#                            optCtrl=list(maxfun=200000),
#                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
                  
#summary(intact_model)
#car::Anova(intact_model)
#
#squashed_model = lmer(data = squashed_data, log10_rt ~ verb + (verb | subject) + 
#                      (1 | stimulus), REML = FALSE, 
 #               lmerControl(optimizer = "bobyqa",
#                            optCtrl=list(maxfun=200000),
#                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
#summary(squashed_model)
#car::Anova(squashed_model)
```

## Sliced model w/ norm covariates
```{r}

sliced_ratings<-read.csv("sliced_covariates.csv")
sliced_acc_data<-exclusions_data%>%filter(state=="slice") %>%left_join(sliced_ratings, by=c("state","stimulus"))

#accuracy
acc_model_sliced_1 = glmer(data = sliced_acc_data, 
                  correct ~ verb+mean_rating + (1|stimulus),
                  family = "binomial", 
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))


acc_model_sliced_2 = glmer(data = sliced_acc_data, 
                  correct ~ verb+mean_rating + (1|subject),
                  family = "binomial", 
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))

summary(acc_model_sliced_2)
anova(acc_model_sliced_1,acc_model_sliced_2)
#2 better fit

#rts
sliced_rt_data<-mad_data%>%filter(state=="slice") %>%left_join(sliced_ratings, by=c("state","stimulus"))

sliced_model = lmer(data = sliced_rt_data, 
                log10_rt ~ verb+mean_rating + (1 + verb|subject) + (1|stimulus),
                REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(sliced_model)
car::Anova(sliced_model)

```
# Demographics and other Descriptives

## Experiment Duration (minutes)
```{r}
duration <- data %>% filter(trial_index == 312) %>%
  mutate(time_elapsed = time_elapsed/1000/60)

duration %>% pull(time_elapsed) %>% hist(40)
duration %>% pull(time_elapsed) %>% boxplot()

duration %>%  summarise(med = median(time_elapsed), iqr = IQR(time_elapsed), max = max(time_elapsed), min = min(time_elapsed))
quantile(duration$time_elapsed, c(.25,.75))
```

## Demographic Info
```{r}
demo = data %>% filter(typeoftrial %in% c("demo", "english") & response != "0") %>%
  separate(response, into = c("demo1", "demo2", "demo3", "demo4"), sep = ":")
age = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo2) %>%
  mutate(age = as.numeric(gsub("[^[:digit:] ]", "", demo2)))
age %>% summarise(mean_age = mean(age))
race = demo %>% filter(demo1 == "{\"Race\"") %>% select(demo2)
race %>% group_by(demo2) %>% count()
gender_education = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo3) %>%
  separate(demo3, into = c("gender", "education"), sep = ",")  %>%
  mutate(gender = tolower(gender),
         gender = gsub("[[:space:]]", "", gender))
gender_education %>% group_by(gender) %>%
  count()
gender_education %>% group_by(education) %>%
  count()
first_lang = demo %>% filter(demo1 == "{\"English\"") %>% select(demo2)
first_lang %>% group_by(demo2) %>% count()
```

```{r}
"18-19" =(age %>% filter(c(age) <= 19 & c(age) >= 18) %>% count())
"20-29" = (age %>% filter(c(age) <= 29 & c(age) >= 20) %>% count())
"30-39" = (age %>% filter(c(age) <= 39 & c(age) >= 30) %>% count())
"40-49" = age %>% filter(c(age) <= 49 & c(age) >= 40) %>% count()
"50-59" = (age %>% filter(c(age) <= 59 & c(age) >= 50) %>% count())
"60+" = (age %>% filter(c(age) >= 60) %>% count())

```


