---
title: "object-state-changes-and squashability"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

# load packages
```{r}
library(tidyverse)
library(ggplot2)
library (ggthemes)
library(rstatix)
library(ggpubr)
library(emmeans)
library(dplyr)
library(lmerTest)
library(paletteer)
library(patchwork)
setwd(here::here())
getwd()
```

# import data
```{r}
prolific1 <- read_csv("../data/prolific_1.csv")
prolific2 <- read_csv("../data/prolific_2.csv")
sona1 <- read_csv("../data/sona.csv")
sona25 <- read_csv("../data/squashability_sona_25.csv")
prolific25 <- read_csv("../data/squashability_prolific_25.csv")
# bind Prolific + Sona
final_data = rbind(prolific1, prolific2, sona1,sona25,prolific25) %>%
  filter(!is.na(subject))

#drop incomplete exp
final_data_filtered <- final_data %>%
  group_by(subject) %>%                  
  filter(n() >= 15) %>%
  ungroup()      

```

# basic descriptives
```{r}
length(final_data_filtered %>% pull(subject) %>% unique())
## N(final) = 275

```
# data treatment

## exclude non-English first language participants***
```{r}
lang <- final_data_filtered %>% filter(typeoftrial == "notenglish") %>% 
  separate(response, into = c("first", "age"), sep = ",") %>% 
  separate(first, into = c("q0", "first"), sep = ":") %>% mutate(first = gsub(" ", "", first)) %>%
  filter(toupper(first) != '"ENGLISH"')

final_data_filtered <- final_data_filtered %>% filter(!subject %in% lang$subject)

length(final_data_filtered %>% pull(subject) %>% unique())

length(lang %>% pull(subject) %>% unique())
```

## exclusions
```{r}
##exclude filler, practice, and non-picture trials
exclusions_data <- final_data_filtered %>%
  filter(typeoftrial=="picture" & !(trialcondition %in% c("filler", "practice"))) %>%
  ##calculate subject accuracy
  mutate(accuracy=ifelse(correct=="TRUE",1,0)) %>% 
  separate(trialcondition, into = c("squashability", "state"))

length(exclusions_data %>% pull(subject) %>% unique())

```

## item labels 
```{r}
items <- gsub('.{0,6}$', '', exclusions_data$stimulus) %>% as.data.frame()
exclusions_data <- exclusions_data %>% mutate(item = items[,1] )

### For experimental picture trials: this code creates a new column 'items' to id the item (picture stimulus) used in each trial. The column 'stimulus' has the picture file name and specifies the smashed/intact picture state. This new col only has the item  (ex. the original stimulus label might be "tomatoS.png" or tomatoN.png" -> new label is "tomato")
```


# accuracy analyses

##accuracy data
```{r}
subject_accuracy <- exclusions_data %>% group_by(subject) %>% summarise(TaskAccuracy=mean(accuracy))
InaccurateSubjects <- subject_accuracy %>% filter(TaskAccuracy < 0.8)

length(InaccurateSubjects %>% pull(subject) %>% unique())
exclusions_data %>% summarise(overall.acc = mean(correct))
```

## plot
```{r}
accuracy_plot<-exclusions_data %>% 
  group_by(squashability, state) %>%
   summarise(ci = list(mean_cl_boot(accuracy) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  mutate(state = fct_recode(state, canonical = "Normal", `non-canonical` = "Smashed" ))%>%
  ggplot(aes(x= state, y = mean, group = squashability, fill = squashability)) +
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "accuracy", title = "")+
  theme_clean()+
  scale_fill_paletteer_d("ggthemes::Classic_10_Light")+
  theme(
    legend.position = "top",
    aspect.ratio = 1.5,
    plot.background = element_rect(color = "white"),
    strip.text.x = element_text(size = rel(2.3)),
    axis.text = element_text(size = rel(2.3)),
    axis.title = element_text(size = rel(3)),
    plot.title = element_text(hjust = 0.5, size = rel(2)),
    legend.text = element_text(size = rel(2)),  
    legend.title = element_text(size = rel(2)))


accuracy_plot
```
## main accuracy model
```{r}
acc_model_0<-glmer(data = exclusions_data, 
                  accuracy ~ state*squashability + (1|subject) + (1|stimulus),
                  family = "binomial", 
                 control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))


acc_model = glmer(data = exclusions_data, 
                  accuracy ~ state*squashability + (1 + state*squashability|subject) + (1|stimulus),
                  family = "binomial", 
                 control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(acc_model)
car::Anova(acc_model)
performance::icc(acc_model)

sjPlot::plot_model(acc_model, type = "int")

anova(acc_model_0,acc_model)
```

Significant interaction

# RT analyses

## rt data
```{r}
##exclude incorrect responses
rt_data <- exclusions_data %>% filter(accuracy==1) %>%
  ##exclude subjects with accuracy < 80%
  filter(!subject %in% InaccurateSubjects$subject)

##total participants after exclusion: N = 237
length(InaccurateSubjects %>% pull(subject) %>% unique())
length(rt_data %>% pull(subject) %>% unique())
```

## transforming RT
```{r}
##make rt numeric
rt_data <- rt_data %>% mutate(rt = as.numeric(rt))
## rt histogram ----does NOT look normally distributed (positive skew)
rt_data %>% filter(rt <10000) %>% 
  ggplot(aes(x=rt)) + geom_histogram(binwidth =20, fill="pink" ) + 
  theme_economist() + labs(x="RT (ms)", y="frequency", title= "RT histogram")

##rt log (base 10) transformation
rt_data <- rt_data %>% mutate(log10_rt=log(rt, 10))

##log10(rt) histogram
rt_data %>% 
  ggplot(aes(x=log10_rt)) + geom_histogram(binwidth =0.1, fill="pink" ) + 
  theme_economist() + labs(x="log10(rt)", y="frequency", title= "RT histogram")
```

## MAD exclusions
```{r}
mad_data = rt_data %>%
  group_by(squashability, state) %>%
  mutate(MAD=3*mad(log10_rt), med=median(log10_rt),
         L= (med - MAD), U=(med + MAD)) %>%
  filter(log10_rt < U) %>% filter(log10_rt > L)

(nrow(rt_data)-nrow(mad_data))/nrow(rt_data)
```

## plot
```{r}
rt_plot<-mad_data %>% 
  group_by(squashability, state) %>%
   summarise(ci = list(mean_cl_boot(rt) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  mutate(state = fct_recode(state, canonical = "Normal", `non-canonical` = "Smashed" ))%>%
  ggplot(aes(x= state, y = mean, group = squashability, fill = squashability)) +
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "response time (ms)", title = "")+
  theme_clean()+
  scale_fill_paletteer_d("ggthemes::Classic_10_Light")+
  theme(
    legend.position = "none",
    aspect.ratio = 1.5,
    plot.background = element_rect(color = "white"),
    strip.text.x = element_text(size = rel(2.3)),
    axis.text = element_text(size = rel(2.3)),
    axis.title = element_text(size = rel(3)),
    plot.title = element_text(hjust = 0.5, size = rel(2)),
    legend.text = element_text(size = rel(2)),  
    legend.title = element_text(size = rel(2))
  )

accuracy_plot+rt_plot
```

## main RT model
```{r}
rt_model = lmer(data = mad_data, 
                log10_rt ~ squashability*state +
                  (1 + squashability+state |subject) + (1|stimulus),
                REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(rt_model)
```


## power analysis
```{r}
library(simr)
#set.seed(123)
#Change the fixed effect to be small
#fixef(rt_model)["squashabilityUnsquashable:stateSmashed"] = 0.05
#Start with an extended model up to 100 participants
#extended_model <- extend(rt_model, along = "subject", n = 200)
#Perform a power curve analysis to test the three-way interaction
#power_curve <- powerCurve(extended_model, 
                         # along = "subject", 
                       # fixed("squashability:state", "anova"), 
                        # nsim = 100)  # Number of simulations per sample size

# Plot the power curve
#plot(power_curve)

#will hit around ~ 225 
```

### assumptions
```{r}
library(glmmTMB)
sjPlot::plot_model(rt_model, type='diag')
sjPlot::plot_model(rt_model, type='resid')
```

### posthoc models 

```{r}
# split into picture type (state)
intact_data = mad_data %>% filter(state == "Normal")
squashed_data = mad_data %>% filter(state == "Smashed")
# run separate models on these data
intact_model = lmer(data = intact_data, log10_rt ~ squashability + (squashability | subject) + 
                      (1 | item), REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
                  
summary(intact_model)
car::Anova(intact_model)
squashed_model = lmer(data = squashed_data, log10_rt ~ squashability + (squashability | subject) + 
                      (1 | item), REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(squashed_model)
car::Anova(squashed_model)
```

Marginal interaction: for "intact", unsquashable RT < squashable RT. For "smashed", squashable RT < unsquashable RT.


## Model w/ norms as continuous predictor
```{r}

squash_ratings<-read.csv("squash_covariates.csv")
squash_acc_data<-exclusions_data %>%left_join(squash_ratings, by="item") %>%mutate(mean_rating=as.numeric(mean_rating))

#accuracy
acc_model_squash = glmer(data = squash_acc_data, 
                  correct ~ state*mean_rating + (state*mean_rating|subject)+(1|stimulus),
                  family = "binomial", 
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl=list(maxfun=200000), 
                      check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))

summary(acc_model_squash)

emmip(
  acc_model_squash,
  state ~ mean_rating,
  at = list(mean_rating = seq(
    min(squash_acc_data$mean_rating, na.rm = TRUE),
    max(squash_acc_data$mean_rating, na.rm = TRUE),
    length.out = 100
  )),
  type = "response",
  CIs = TRUE
)

#rts
squash_rt_data<-mad_data%>%left_join(squash_ratings, by="item")

squash_model = lmer(data = squash_rt_data, 
                log10_rt ~ state*mean_rating + (state*mean_rating|subject) + (1|stimulus),
                REML = FALSE, 
                lmerControl(optimizer = "bobyqa",
                            optCtrl=list(maxfun=200000),
                    check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))
summary(squash_model)
car::Anova(squash_model)

```

# Demographics and other Descriptives

## Experiment Duration (minutes)
```{r}
duration <- final_data %>% filter(typeoftrial == "english") %>%
  mutate(time_elapsed = time_elapsed/1000/60)

duration %>% pull(time_elapsed) %>% hist(40)
duration %>% pull(time_elapsed) %>% boxplot()

duration %>%  summarise(med = median(time_elapsed), iqr = IQR(time_elapsed), max = max(time_elapsed), min = min(time_elapsed))
quantile(duration$time_elapsed, c(.25,.75))
```

## Demographic Info
```{r}
demo = final_data %>% filter(typeoftrial %in% c("demo", "english") & response != "0") %>%
  separate(response, into = c("demo1", "demo2", "demo3", "demo4"), sep = ":")
age = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo2) %>%
  mutate(age = as.numeric(gsub("[^[:digit:] ]", "", demo2)))
age %>% summarise(mean_age = mean(age))
race = demo %>% filter(demo1 == "{\"Race\"") %>% select(demo2)
race %>% group_by(demo2) %>% count()
gender_education = demo %>% filter(demo1 == "{\"Age\"") %>% select(demo3) %>%
  separate(demo3, into = c("gender", "education"), sep = ",")  %>%
  mutate(gender = tolower(gender),
         gender = gsub("[[:space:]]", "", gender))
gender_education %>% group_by(gender) %>%
  count()
gender_education %>% group_by(education) %>%
  count()
first_lang = demo %>% filter(demo1 == "{\"English\"") %>% select(demo2)
first_lang %>% group_by(demo2) %>% count()
```

```{r}
"18-19" =(age %>% filter(c(age) <= 19 & c(age) >= 18) %>% count())
"20-29" = (age %>% filter(c(age) <= 29 & c(age) >= 20) %>% count())
"30-39" = (age %>% filter(c(age) <= 39 & c(age) >= 30) %>% count())
"40-49" = age %>% filter(c(age) <= 49 & c(age) >= 40) %>% count()
"50-59" = (age %>% filter(c(age) <= 59 & c(age) >= 50) %>% count())
"60+" = (age %>% filter(c(age) >= 60) %>% count())

```

# Normative analyses of stimuli

## import data and organize
```{r}
data <- read_csv("../data/survey_data.csv") %>% filter(typeoftrial == "survey") %>% 
  separate(trialcondition, into =c("condition", "item"), sep="_") %>%
  separate(response, into = c("label", "response"), sep = ":") %>%
  mutate(response=parse_number(response)) %>%
  mutate(rt = as.numeric(rt))%>%
  mutate("logRT" = log10(rt)) %>%
  mutate(item = as.character(item))%>%
  mutate(condition = fct_recode(condition, squashable = "s", filler = "f", unsquashable="u"),
         condition = fct_relevel(condition,  "unsquashable","filler", "squashable"))

length(data %>% pull(subject) %>% unique())

#get covariate values
#squash_covariates <- data %>%
#  mutate(item = as.numeric(item)) %>%
#  filter(!is.na(item)) %>%
#  group_by(item,condition) %>%
#  summarise(
#    mean_rating = mean(response, na.rm = TRUE),
#    .groups = "drop"
#  ) %>%
#  arrange(item)

#write.csv(squash_covariates, "squash_covariates.csv")
```

## basic plot
```{r}
id_ratings = data %>% group_by(subject, condition) %>%
  summarise(mean_rating = mean(response))
data %>% 
  group_by(condition) %>%
  summarise(ci = list(mean_cl_boot(response) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% 
  unnest %>%
  ggplot(aes(x=condition, y=mean)) +
  geom_bar(aes(fill = condition), stat="identity", position="dodge", width=0.7, color="gray24") +
  geom_errorbar(aes(ymin = lwr, ymax = upr), size = 0.5, width = 0.15, 
                color = "black", position = position_dodge(0.70)) +
  labs(y = "mean ratings", title = "") +
  theme_clean() +
  scale_fill_paletteer_d("ggthemes::Classic_10_Light") +
  theme(plot.background = element_rect(color = "white"),
        strip.text.x = element_text(size = rel(2)),
        axis.text = element_text(size = rel(1.5)),
        axis.title = element_text(size = rel(2)),
        plot.title = element_text(hjust = .5, size = rel(1.5)),
        legend.position = "none")
```

# model
```{r}
library(ordinal)
# make response a factor
data = data %>% mutate(ordinal_response = as.factor(response))
# set the squashable category as a reference category for the model
contrasts(data$condition) = contr.treatment(3, base = 2)
# run an ordinal mixed effects regression
mod = clmm(ordinal_response~condition+(1|subject), data=data, Hess=T)
summary(mod)
```

