---
title: "Object State Sanity Check"
author: "Channing Hambric"
date: "2024-12-02"
output: html_document
---


# load packages
```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
library(dplyr)
library(stringr)
```

# import data
```{r}
#Data from third self pilot - N=8
check_data <-read.csv("../data/object-state-sentence-focus_checking4.csv", fileEncoding = "UTF-8")

base<-check_data
```

#participant counts
```{r}
check_data_clean <- check_data %>% 
  filter(!is.na(subject))

length(check_data_clean %>% pull(subject) %>% unique())

#Should be 8 participants
```
#DESIGN:

Practice - 4 trials total, 2 y, 2 n, 2 in target format 2 in filler format

Experimental Trials - 40 target trials, 48 filler trials

Of target trials, 5 trials each in the following conditions:
Heavy_Normal_Focus
Heavy_Smashed_Focus
Heavy_Normal_NotFocus
Heavy_Smashed_NotFocus
Light_Normal_Focus
Light_Smashed_Focus
Light_Normal_NotFocus
Light_Smashed_NotFocus

Half of filler (24) get comprehension questions (12 y, 12 n)

#checking practice
```{r}
prac_data <- check_data %>%
  filter(trialcondition %in% c("practice_y","practice_n")) %>% filter(correct!="NA")

#Each participant should get 4 practice, 2 with yes (true) as the correct answer and 2 with no (False)
prac_data %>% group_by(subject,trialcondition) %>% count()

#Practice sentences should have 2 framed as targets ("came into contact with"), 2 framed as filler ("x is on a x")
prac_data %>% select(sentence) %>% unique()

```


#checking experimental trials
```{r}
exp_data <- check_data %>%
  filter(typeoftrial=="picture" & !(trialcondition %in% c("filler", "practice_y","practice_n"))) %>%
  separate(trialcondition, into = c("weight", "state","focus"))

#There should be 320 trials total (8 sub x 40)
exp_data %>% count()

#There should be 40 trials per participant
exp_data %>% group_by(subject) %>% count()

#Each participant should have 20 trials in each weight condition
exp_data %>% group_by(subject, weight) %>% count()

#Each participant should have 20 trials in each state condition
exp_data %>% group_by(subject, state) %>% count()

#Each participant should have 20 trials in each focus condition
exp_data %>% group_by(subject, focus) %>% count()

#Each participant should have 5 trials in each of 8 weight x state x focus conditions
exp_data %>% group_by(subject, weight, state, focus) %>% count()

#Across participants, there should be 40 trials in each weight x state x focus condition (8 sub x 5 trials)
exp_data %>% group_by(weight, state, focus) %>% count()


#Commenting out since not all conditions sampled
#Each item should appear in each  of 8 conditions across lists

items <- gsub('.{0,5}$', '', exp_data$object) %>% as.data.frame()
exp_data <- exp_data %>% mutate(item = items[,1] )

exp_data %>%
  select(item, weight, state, focus) %>% unique() %>%
  group_by(item) %>% arrange(item) %>%
  summarise(count = n()) %>%  
  mutate(status = case_when(
    count < 8 ~ "item does not appear in all conditions",
    count == 8 ~ "item appears in all conditions"
  ))


#Checking that list corresponds to condition
exp_data %>%select(condition,List) %>%unique()

#Checking that all normal/smashed pictures were labeled as such
lab_data <- exp_data %>%
  mutate(check_state = str_sub(object, -5), 
         label = case_when(
           (check_state %in% c("S.jpg","S.png") & state == "Smashed") ~ TRUE,  
           (check_state %in% c("N.jpg","N.png") & state == "Normal") ~ TRUE,  
           TRUE ~ FALSE  
         )) %>% select(check_state,object,label)

if (all(lab_data$label == TRUE)) {
  print("all images labeled correctly")
} else {
  print("Some images mislabeled")
}

#Checking if weight labeled correctly 
weight_data <- exp_data %>%
  mutate(weight_label = case_when(
           str_detect(sentence, "brick") & weight == "Heavy" ~ TRUE,  
           str_detect(sentence, "feather") & weight == "Light" ~ TRUE, 
           TRUE ~ FALSE 
         )) %>%
  select(weight, sentence, weight_label)

if (all(weight_data$weight_label == TRUE)) {
  print("all images labeled correctly")
} else {
  print("Some images mislabeled")
}

#Checking focus labeling
exp_data <- exp_data %>%
  mutate(item = case_when(
    item == "tortita" ~ "rice cake",         
    item == "lamp" ~ "light bulb",          
    item == "can" ~ "can of beer",
    item == "icecream" ~ "ice cream bar",          
    TRUE ~ item                             
  ))
focus_data <- exp_data %>%
  mutate(focus_label = case_when(
           str_detect(str_remove_all(str_to_lower(str_remove_all(sentence, "[-[:space:]]")), "\\.$"), 
                      paste0(str_to_lower(str_remove_all(item, "[-[:space:]]")), "$")) & focus == "NotFocus" ~ TRUE,
           !str_detect(str_remove_all(str_to_lower(str_remove_all(sentence, "[-[:space:]]")), "\\.$"), 
                        paste0(str_to_lower(str_remove_all(item, "[-[:space:]]")), "$")) & focus == "Focus" ~ TRUE,  
           TRUE ~ FALSE 
         )) %>%
  select(item,sentence, focus, focus_label)

#Note that item/.jpg labels differed for lamp (light bulb), tortita (rice cake), beer can (can), and ice cream bar (ice cream)

if (all(focus_data$focus_label == TRUE)) {
  print("all images labeled correctly")
} else {
  print("Some images mislabeled")
}


```

#checking filler
```{r}
filler_data <- check_data %>%
  filter(trialcondition=="filler") %>% filter(index!="NA")

#There should be 384 total (8 sub x 48 trials)
filler_data %>% count()

#There should be 48 trials per participant
filler_data %>% group_by(subject) %>% count()

#Should be 192 comprehension q's (24 per 8 sub)
check_data %>% filter(typeoftrial=="comprehension") %>% count()

#Should be 24 comprehension q's per sub
check_data %>% filter(typeoftrial=="comprehension") %>% group_by(subject) %>%count()

#Should be 12 yes and 12 no comp q's per subject
check_data %>% filter(typeoftrial=="comprehension") %>% group_by(subject,comprehension_type) %>%count()

```

#Checking Demos (are req q's answered)
```{r}
#check that all p's have an age, gender, and education values

demo = base %>% filter(typeoftrial %in% c("demo","english","notenglish")) 

#Should be 8 age entries
demo %>% select(subject,age) %>%na.omit() %>%count()

#Should be 8 gender entries
demo %>%select(subject, gender) %>%       
  filter(!is.na(gender) & gender != "") %>%  
  count() 

#Should be 8 ed entries
demo %>% select(subject,education) %>%na.omit() %>%count()


#Check count of non/native speakers
#0 = native speaker
demo %>% filter(typeoftrial=="english") %>%group_by(response) %>%count()

#Getting English AOA
demo_sep <- base %>% filter(typeoftrial %in% c("notenglish")) %>%
  separate(response, into = c("Not Native Speaker","AOA"), sep = ",")%>%
  mutate(AOA = gsub("[^0-9]", "", AOA)) %>%  # Keep only numeric values in AOA
  filter(AOA != "")  # Remove rows where AOA is empty
demo_sep %>%select(subject,AOA)



```





