---
title: "R Notebook"
output: html_notebook
---
# load packages
```{r}
library(tidyverse)
library(ggplot2)
library (ggthemes)
library(rstatix)
library(ggpubr)
library(emmeans)
library(dplyr)
```

# import data and organize
```{r}
data <- read_csv("survey_data.csv") %>% filter(typeoftrial == "survey") %>% 
  separate(trialcondition, into =c("condition", "item"), sep="_") %>%
  separate(response, into = c("label", "response"), sep = ":") %>%
  mutate(response=parse_number(response)) %>%
  mutate(rt = as.numeric(rt))%>%
  mutate("logRT" = log10(rt)) %>%
  mutate(item = as.character(item))%>%
  mutate(condition = fct_recode(condition, squashable = "s", filler = "f", unsquashable="u"),
         condition = fct_relevel(condition,  "unsquashable","filler", "squashable"))
```

# basic plot
```{r}
id_ratings = data %>% group_by(subject, condition) %>%
  summarise(mean_rating = mean(response))
data %>% 
  group_by(condition) %>%
  summarise(ci = list(mean_cl_boot(response) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% unnest%>%
  ggplot(aes(x= condition, y = mean)) +
geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "gray24")+
  geom_point(data = id_ratings, aes(x = condition, y = mean_rating),
             shape = 21, position = position_jitter(width = 0.05),
             alpha = 0.3)+
  geom_errorbar(aes(ymin=lwr, ymax=upr), size = 0.5, width=.15, 
                color = "black", position = position_dodge(0.70))+
  labs(y = "mean ratings", title = "squashability ratings")+
  theme_clean()+
  scale_fill_manual(values = c("darksalmon", "darkslategray4","darkolivegreen4"))+
  theme(plot.background = element_rect(
    color = "white"),
    strip.text.x = element_text(size =rel(2)),
        axis.text = element_text(size =rel(1.5)),
        axis.title = element_text(size =rel(2)),
        plot.title = element_text(hjust = .5, size = rel(1.5)),
        legend.position = "none")
```

# model
```{r}
library(ordinal)
# make response a factor
data = data %>% mutate(ordinal_response = as.factor(response))
# set the squashable category as a reference category for the model
contrasts(data$condition) = contr.treatment(3, base = 2)
# run an ordinal mixed effects regression
mod = clmm(ordinal_response~condition+(1|subject), data=data, Hess=T)
summary(mod)